The Power of Two Random Choices

### 2014-01-13 article/article.tex
fix: citations in ascending order
Application: Low congestions circuit routing

It is easy to see how the balls-into-bins idea can be applied: Assuming a perfect hash function\footnote{i.e., each of the $n$ elements is mapped to one of the $n$ table entries uniformly and independently at random}, the elements are balls to be mapped to table entries or bins, implying a worst case look-up time of $\Theta\left(\frac{\log n}{\log \log n}\right)$ w.h.p. Applying the two-choice paradigm by using two hash functions and mapping elements to the least full of two table entries implies a bound of $\Theta\left(\log \log n\right)$ w.h.p. This approach is easy to parallelize, it does not require re-hashing of data and requires only two hash functions \cite{MRS01} \cite{ABKU99}.

@@

To maximize the throughput of the servers, a uniform distribution of tasks to servers is desireable. This can easily be achieved by a central dispatcher. However, such a central dispatcher poses a bottleneck within a distributed system and querying $m$ servers for their current load involves significant overhead, e.g., sending a message to each server and waiting for the reply and more efficient solutions are neccessary. As with hashing, the balls-into-bins model can be applied easily: we need to assign tasks (the balls) to servers (the bins). Assuming tasks are assigned to servers uniformly and independently at random, the two-choice paradigm is directly applicable: If each client samples two random servers for their load and sends issues its request to the least loaded server, the overhead is comperatively small and the difference in load on the $n$ servers is bounded by $\Theta\left(\log \log n\right)$ w.h.p. \cite{RS98} \cite{MRS01} \cite{KLM92}.

\subsection{Dynamic resource allocation}

@@

>>>>

It is easy to see how the balls-into-bins idea can be applied: Assuming a perfect hash function\footnote{i.e., each of the $n$ elements is mapped to one of the $n$ table entries uniformly and independently at random}, the elements are balls to be mapped to table entries or bins, implying a worst case look-up time of $\Theta\left(\frac{\log n}{\log \log n}\right)$ w.h.p. Applying the two-choice paradigm by using two hash functions and mapping elements to the least full of two table entries implies a bound of $\Theta\left(\log \log n\right)$ w.h.p. This approach is easy to parallelize, it does not require re-hashing of data and requires only two hash functions \cite{ABKU99} \cite{MRS01}.

@@

To maximize the throughput of the servers, a uniform distribution of tasks to servers is desireable. This can easily be achieved by a central dispatcher. However, such a central dispatcher poses a bottleneck within a distributed system and querying $m$ servers for their current load involves significant overhead, e.g., sending a message to each server and waiting for the reply and more efficient solutions are neccessary. As with hashing, the balls-into-bins model can be applied easily: we need to assign tasks (the balls) to servers (the bins). Assuming tasks are assigned to servers uniformly and independently at random, the two-choice paradigm is directly applicable: If each client samples two random servers for their load and sends issues its request to the least loaded server, the overhead is comperatively small and the difference in load on the $n$ servers is bounded by $\Theta\left(\log \log n\right)$ w.h.p. \cite{KLM92} \cite{RS98} \cite{MRS01}.

@@

\subsection{Low congestion circuit routing}
In multi-stage interconnection networks, virtual circuit-switching is used to route communication. For simplicity, we consider the \emph{permutation routing problem}, where one request originates at each of the $n$ input nodes and one request is destined for each of the $n$ output nodes and assume a \emph{butterfly network} as the underlying interconnection network (see \cite{CLR09} for a detailed definition). Then, congestions is the maximum number of paths that are routed through a network link. The goal of a \emph{circuit routing algorithm} is to to allocate paths for all $n$ communication requests such that congestion is minimal.

In this context, \emph{Valiant's paradigm} \cite{V82} states that any such permutation routing problem can be solved by transforming it into two problems: First, find a path for each request to an intermediate destination chosen uniformly and independently at random. Next, find a path from the intermediate to the actual destination. This routing technique is analogous to the single-choice balls-into-bins problem where each communication request (or ball) chooses an intermediate destination (a bin) at random. In this analogy, the congestion achieved by Valiant's paradigm corresponds to the maximum load in the balls-into-bins model, thus implying a bound of $\Theta\left(\log \log n\right)$ on the congestion w.h.p. \cite{CMM+98} \cite{MRS01}.

Applying the two-choice paradigm in this context is somewhat more involved and beyond the scope of this paper. The interested reader may find more details in \cite{CMM+98}.
